---
title: "Time series analysis project"
author: "Sofia Davoli"
date: "3/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE)
library("tidyverse")
library("xts")
library("forecast")
library("KFAS")
library("lubridate")
library("caret")
library("astsa")
library("urca")
library("tsfknn")
library("ModelMetrics")
library('tensorflow')
library("keras")
library("DMwR")
library("ggpubr")


load("C:/Users/davol/Documents/Magistrale/time series/consegna/SDMSTA_813479_obj.RData")
#load("C:\\Users\\davol\\Documents\\Magistrale\\time #series\\PROJECT\\object\\obj_sdmtsa.RData")
#load("C:\\Users\\davol\\Documents\\Magistrale\\time #series\\PROJECT\\obj_UCM.RData")
#load("C:\\Users\\davol\\Documents\\Magistrale\\time #series\\PROJECT\\obj_KNN.RData")

ts <- read_csv2("C:\\Users\\davol\\Documents\\Magistrale\\time series\\PROJECT\\TrainingSet.csv")
ts <- ts %>%
  rename(Data = DATA,
        Valore = VALORE)
```

## Data Preprocessing and Exploration

Data preprocessing consist in fixing 2 missing value in this time series (TS). Missing values are on 29-03-2020 and 31-03-2019 which represents Italian daylight savings time (ora legale). Missing imputation has been applied by substituting the value with the one of the previous hour, considering that values at 3am seems to be quite constant.

For further exploration, weekly TS has been created by setting frequency parameter to 168 (24 hours* 7 days).

Finally TS has been scaled to optimize models training.


```{r data preprocessing, results='hide'}

#head(ts)

#-- Explorative Analysis
sapply(ts, function(x)(sum(is.na(x)))) #-- find NA

ts %>%
  group_by(Ora) %>% #-- Group by Ora
  summarise(n()) #-- find NA (problem with 3 A.M.)

ts %>%
  group_by(month(date(Data)),
           day(date(Data))) %>%
  summarise(n()) %>%
  filter(`n()` == 47) #-- find days with legal hours

#-- problem with ora legale: 31-03-2019 and 29-03-2020
#-- add in the ts
ts_legale <- tribble(
  ~Data, ~Ora, ~Valore,
  "2019-03-31", 3, 3039997,
  "2020-03-29", 3, 2329514
) %>%
  mutate(Data = as.Date(Data))

#-- Weekly Time series
ts_weekly <- ts %>%
  bind_rows(ts_legale) %>% #-- Concatenazione
  arrange(Data) %>% #-- Ordinamento
  ts(frequency = 168) #-- Frequenza Settimanale (24 h * 7 g)

#-- Daily Time series
ts_daily <- ts %>%
  bind_rows(ts_legale) %>%
  arrange(Data) %>%
  ts(frequency = 24) #-- 24*7

#-- Time Series
ts <- ts %>%
  bind_rows(ts_legale) %>%
  arrange(Data)

#-- identify min max values
max_value <- which.max(ts$Valore)
min_value <- which.min(ts$Valore)

#-- Mean Aggregaton for Seasonal Plot
ag <- aggregate(Valore ~ Data, ts, mean)

#-- scaling TS to optimize models training
ts_stand <- ts(scale(ts$Valore), frequency = 24)
```

Daily plot (1) shows a clear daily seasonality but from Weekly plot (2) no particular weekly trend appears, in fact there are no significant differences in values from 1st day of the week from the last day. We can assume that weekend days have no significant influence  on TS. 
Daily seasonality is confirmed by plot 6 (automatic TS decomposition) in which appear a seasonal24 decomposition. 


```{r daily and weekly seasonality , echo = FALSE, warning=FALSE, cache = TRUE}
#-- daily Plot
ggseasonplot(ts_daily[,"Valore"],
             year.labels=TRUE,
             year.labels.left=FALSE,
             ylab = "Value (in millions)",
             xlab = "Day Hours",
             main = "1. Daily plot") +
  theme_bw() +
  theme(axis.text.y = element_text(face = "bold"),
        axis.text.x = element_text(face = "bold",
                                   angle = 45)) + 
  scale_y_continuous(breaks = c(2000000,
                                4000000,
                                6000000,
                                8000000),
                     labels = c(2,4,6,8)) +
  scale_x_continuous(breaks = c(2, 4, 6),
                     label = c(2,3,4))+theme_gray()
#-- weekly Plot
ggseasonplot(ts_weekly[,"Valore"], year.labels=TRUE, 
             year.labels.left=TRUE, 
             ylab = "Value (in millions)",
             xlab = "Day of the week",
             main = "2. Weekly plot")+theme_gray()

```

Monthly plot clearly shows no monthly seasonality.
Finally a clear yearly seasonality appear in plot 4 (even if only 2 year are available). In this plot line 2 is always below line 1 because of Covid effect, this effect can easily be seen from total time series plot (5) where the second half of the plot decrease compared from first half (red line indicate the trend and it decrease from one year to the other).




```{r monthly and yearly seasonality , echo = FALSE, cache = TRUE}
#-- monthly Plot
ggseasonplot(ts(ag[,"Valore"], frequency = 30), year.labels=TRUE,                year.labels.left=TRUE,
             ylab = "Value (in millions)",
             xlab = "days of the month",
             main = "3. Monthly plot")+theme_gray()
#-- yearly Plot
ggseasonplot(ts(ag[,"Valore"], frequency = 730/2),       year.labels=FALSE, year.labels.left=FALSE, labelgap = 50,
             ylab = "Value (in millions)",
             xlab = "Months", 
             main = "4. Yearly plot")+theme_gray()
#-- ts plot
ts %>%
  ggplot(aes(x=Data, y=Valore)) +
      ggtitle("5. TS plot") +
      xlab("time") + 
      ylab("Value (in millions)")+
      geom_line() +
  stat_smooth(method = "gam", se = T, col="blue", level=0.99) +
  stat_smooth(method = "lm", se = T, col="red", level=0.99)+theme_gray()

#-- scomposizione serie storica
autoplot(mstl(ts(ts$Valore, frequency = 24)), main='6. Decomposition of TS')+theme_gray()

```




## Arima and Naive methods

Training and validation set for arima models are created respectively using  80% - 20% of data. 
Models are trined and evaluated on scaled data. 

```{r training-validation arima, echo = TRUE, results='hide', cache = TRUE}
train_arima <- ts(ts_stand[1:as.integer(nrow(ts)*0.8), ],
                  frequency = 24)
validation_arima <- ts(ts_stand[(1 + as.integer(nrow(ts)*0.8)) :                      nrow(ts), ], frequency = 24)

```

Some preliminary assumptions are made based on ACF and PACF results.

ACF plot shows a clear seasonality of period 24, in fact 
the autocorrelations is higher for the seasonal lags (at multiples of 24) than for other lags.
PACF plot suggest an AR(2),AR(3) or AR(5) model since the number of non-zero partial autocorrelations (in this case 5) gives the order of the AR model.


```{r acf and pacf plot}
#--- Acf Pacf 
ggarrange(ggAcf(ts_stand,
      lag.max = 48, #-- 2 Days for Seasonality
      main = "ACF"),
          ggPacf(ts_stand,
       lag.max = 48, #-- 2 Days for Seasonality
       main = "PACF"),
          nrow = 2, ncol= 1)+theme_gray()


```


To identify the best model Box-jenkins methodology is applied.

1 - BoxCox optimal transformation suggest as optimum lambda 0.9284836 which is very close to 1, meaning that no other trasformation is needed to stabilize variance.

2 - As previously seen, the TS show a strong seasonality component. To remove it, a seasonal differences at lag 24 is taken. A Box-Ljung test reject H0 of White Noise. 

3 - 2 more differences has to be taken to get to a White noise process. In first differences ACF and PACF show a wery high first lag, suggesting that another differences has to be taken. ACF plot of second differences suggest a White noise process confirmed by Box-Ljung test

This results suggest the use of a SARIMA model with daily seasonality, 2 integration and AR component equal to 2, 3 or 5.

```{r Box-jenkins methodology, echo = FALSE, cache = TRUE }
#check variance
lambda <- BoxCox.lambda(train_arima)
#lambda

#tseasonal differences
train_diff_seasonal<-diff(train_arima, lag=24)
autoplot(train_diff_seasonal, main='seasonal difference')+theme_gray()
Box.test(train_diff_seasonal, type='Ljung-Box') #H0 rejected

#first differences
train_diff_trend<-diff(train_diff_seasonal, differences = 1)
autoplot(train_diff_trend, main='first differences')+theme_gray()#c'e ancora stagionalità giornaliera
ggarrange(ggAcf(train_diff_trend, 
                lag.max = 72, main = "ACF"),
          ggPacf(train_diff_trend, 
                 lag.max = 72, main = "PACF"),
          nrow = 2, ncol= 1)+theme_gray()



Box.test(train_diff_trend, type='Ljung-Box') #H0 rejected

#second differences
train_diff_trend2<-diff(train_diff_trend, differences = 1)
autoplot(train_diff_trend2, main='second difference')+theme_gray()
ggAcf(train_diff_trend2, lag.max = 72, main = "ACF after 1 seasonal difference and 2 difference")+theme_gray()
Box.test(train_diff_trend2, type='Ljung-Box') #H0 accepted

```


A naive model is trained. Comparing plot 7 (showing predicted values of naive model) with the real TS plot is clear that this model doesn't show any yearly seasonality component. ACF and PACF of residuals are not white noises typical.  Mae of naive model on validation data is 0.6563888. 

```{r naive model, echo = FALSE, cache = TRUE}
naive <- naive(train_arima, h = 3504)
autoplot(naive, xlab = "Time", ylab = "Value",  main = "7. Naive model prediction" ) 
naive_res<-naive$residuals

ggarrange(ggAcf(naive_res,
                lag.max = 96,       
                main = ""),
          ggPacf(naive_res,
                 lag.max = 96,       #-- 2 Days for Seasonality
                 main = ""),
          labels = c("ACF naive", "PACF naive"),
          nrow = 2, ncol= 1)+theme_gray()

Box.test(naive_res, type='Ljung-Box') #H0 rejected

naive_mean <- ts(naive$mean, start = 1, frequency = 24)
mae_naiff <- mean(abs(validation_arima - naive_mean))
#mae_naiff

```

First Arima trained has AR(2),I(1) and MA(1). 

Residual analysis suggest to reject H0 of white noise. Moreover in Acf plot a strong seasonal component is present (increase MA order) and an increase in AR part of the model is suggested from PACF.
Mae of ARIMA(2,1,1) on validation data is 0.5436. 

```{r arima(2,0,0)(0,1,1), cache=TRUE}
mod1_arima <- Arima(train_arima, c(2,0,0), c(0,1,1), lambda = "auto",
              include.constant = TRUE)
smod1_arima <- summary(mod1_arima)
#smod1_arima[,"MAE"]

checkresiduals(mod1_arima)
ggPacf(mod1_arima$residuals, lag.max = 96, main = "PACF")+theme_gray()
#Box.test(mod1_arima$residuals, type='Ljung-Box') #H0 rejected
#-- previsioni
pred_val1 <- forecast(mod1_arima, h = 3504)

mae_mod1_arima <- mean(abs(ts(pred_val1$mean, start = 1, frequency = 24) - validation_arima))
#mae_mod1_arima


```

Arima (3,1,2) is trained. 

Residual analysis suggest to accept H0 of white noise, it looks like this model is a good model. Moreover Acf and Pacf are typical of white noises. 
Prediction in plot 8 are not as good as expected, in particuar the last part of prediction sottoestimate the real TS values and it looks like the arima model perceive a decreasing trend, surely caused by Covid effect. Arima are not very good for long time predictions. 
Mae of ARIMA(3,1,2) on validation data is 0.5206016. 


```{r arima(3,0,0)(0,1,2), cache=TRUE}

mod2_arima <- Arima(train_arima, c(3,0,0),                    seasonal=list(order=c(0,1,1),period=24),
                    lambda = "auto",
                    include.constant = TRUE)
smod2_arima <- summary(mod2_arima)
#smod2_arima[,"MAE"]

ggarrange(ggAcf(mod2_arima$residuals,
                lag.max = 96,       
                main = ""),
          ggPacf(mod2_arima$residuals,
                 lag.max = 96,       #-- 2 Days for Seasonality
                 main = ""),
          labels = c("ACF arima(3,0,0)(0,1,2)", "PACF arima(3,0,0)(0,1,2)"),
          nrow = 2, ncol= 1)
#checkresiduals(mod2_arima)
Box.test(mod2_arima$residuals, type='Ljung-Box') #H0 rejected

#-- previsioni
pred_val2 <- forecast(mod2_arima, h = 3504)

valid <- ts(ts_stand[(1 + as.integer(nrow(ts)*0.8)) : nrow(ts), ]
            , frequency = 24, start=583) #per plot

autoplot(train_arima, main='8. Arima (3,1,2)  prevision')+autolayer(valid)+autolayer(pred_val2$mean)+theme_gray()

mae_mod2_arima <- mean(abs(ts(pred_val2$mean, start = 1, frequency = 24) - validation_arima))
#mae_mod2_arima

```

Auto Arima model is trained.

Residual analysis suggest to accept H0 of white noise, it looks like this model is a good model. Moreover Acf and Pacf are typical of white noises.
Also this model sottoestimate the second half of validation data. 
Mae of Auto Arima on validation data is 0.496358, 0.03 less than the previous simplest model. 
Arima are not very good for long time predictions, in fact R. Hyndman explain that auto.arima function and arima model in general are designed for short forecasting period. He sugggest to use a Fourier series approach where the seasonal pattern is modelled using Fourier terms with short-term time series dynamics allowed in the error.

REF: https://robjhyndman.com/hyndsight/longseasonality/


```{r ARIMA(5,0,2)(2,1,0)[24], cache=TRUE}
automod_arima <- auto.arima(train_arima, lambda = "auto")
sautomod_arima <- summary(automod_arima)
#sautomod_arima[, "MAE"]

checkresiduals(automod_arima)
#Box.test(automod_arima$residuals, type='Ljung-Box') #H0 accepted
ggarrange(ggAcf(automod_arima$residuals,
                lag.max = 96,       
                main = ""),
          ggPacf(automod_arima$residuals,
                 lag.max = 96,       #-- 2 Days for Seasonality
                 main = ""),
          labels = c("ACF auto.arima", "PACF auto.arima"),
          nrow = 2, ncol= 1)


pred_val_auto <- forecast(automod_arima, h = 3504)

mae_automod_arima <- mean(abs(ts(pred_val_auto$mean, start = 1, frequency = 24) - validation_arima))
#mae_automod_arima


autoplot(train_arima, main='9. Auto Arima model fitting and prevision')+autolayer(pred_val_auto$fitted)+autolayer(valid)+autolayer(pred_val_auto$mean)


```

Arimax models are trained. 

Function mtst is used to create a multiseasonal (daily and yearly) time series objects. Moreover 2 fourier regressor are taken to handle the daily seasonality and 2 regressor for yearly seasonality.
Both arima(3,1,2) and auto.arima model are trained using fourier regressors.
Best results are obtained using auto arimax with a MAE of 0.4367472.
From residuals analysis White noise assumption is accepted.

```{r ARIMAX, cache=TRUE}
y<-msts(train_arima, c(24, 365*24)) # multiseasonal ts

#-- arimax
arimax_1 <- Arima(y, c(3,1,2), xreg=fourier(y, K=c(2,2)))

#-- previsioni e mae
pred_arimax_1 <- forecast(arimax_1, xreg= fourier(y, K=c(2,2), 3504), 3504)
plot(pred_arimax_1, main='10. Arimax(3,1,2) prevision' )

mae_arimax1 <- mean(abs(ts(pred_arimax_1$mean, start = 1, frequency = 24) - validation_arima))
#mae_arimax1

checkresiduals(arimax_1)

#-- auto arimax
arimax_2 <- auto.arima(y, seasonal=F, xreg=fourier(y, K=c(2,2)))


#-- previsioni e mae
pred_arimax_2 <- forecast(arimax_2, xreg= fourier(y, K=c(2,2), 3504), 3504)
plot(pred_arimax_2, main='11. Auto Arimax prevision' )


mae_arimax_2 <- mean(abs(ts(pred_arimax_2$mean, start = 1, frequency = 24) - validation_arima))
#mae_arimax_2 

checkresiduals(arimax_2)

```

Arima (3,1,2) with fourier regressor is chosen as best model because:

 1 residuals are white noises
 2 increase of 2 in the AR part only generate 0.01 of decreasing in MAE, so a simplest model is preferred.


```{r test prevision  with best ARIMA}

ts_best<-msts(ts_stand, c(24, 365*24))
best_arima <- Arima(ts_best,
                         c(3,1,2),
                         xreg=fourier(ts_best, K=c(2,2)),
                    include.constant = TRUE,
                    lambda = 'auto')
                    
sbest_arima <- summary(best_arima)
#sbest_arima[,"MAE"]

pred_best_arima <- forecast(best_arima, xreg= fourier(ts_best, K=c(2,2), 1464), 1464)
autoplot(pred_best_arima, main='12. Test prevision best arimax', ylab='standardized TS')



```

## UCM 

Training and validation set for ucm models are created respectively using  80% - 20% of data. 
Models are trined and evaluated on scaled data. 
Variance of training set is extracted to initialize optimization parameters.

```{r training-validation UCM,  echo = TRUE}
train_ucm <- ts(ts_stand[1:as.integer(nrow(ts_stand)*0.8), ],
                frequency = 24,
                start = 1)

validation_ucm <- ts(ts_stand[(1 + as.integer(nrow(ts_stand)*0.8)) : nrow(ts_stand), ],
                     frequency = 24,
                     start = 585)

v_train_ucm <- var(train_ucm) 
```

Firs UCM trained is a simple Local Linear Trend, MAE obtained is 0.6857667. Plot 13 show prevision made with this model. 

```{r Local Linear Trend}
mod1_ucm <- SSModel(train_ucm ~ SSMtrend(2, 
                Q = list(matrix(NA),matrix(0))), 
                H = matrix(NA)) 


fit1_ucm <- fitSSM(mod1_ucm,
                   inits = log(c(v_train_ucm/10,
                                 v_train_ucm/20,
                                 v_train_ucm/100))) 
#fit1_ucm$optim.out$convergence #-- convergence

#-- prevision
pred1_ucm <- predict(fit1_ucm$model,
                     n.ahead = length(validation_ucm)) 
pred1_ucm <- ts(pred1_ucm, start = 585, frequency = 24)


plot(validation_ucm, main = "13. LLT previsions on validation", type = "l")
lines(pred1_ucm, col = "red")

mae1_ucm <- mean(abs(validation_ucm - pred1_ucm))
#mae1_ucm
```

LLT with daily stochastic dummy.

Seasonal dummys are used to insert in the model daily seasonality. As seen before it is necessary to model this seasonality which represent a strong component of this TS. With this model a MAE of 0.511847 is obtained.
Focus of prevision made in first and last week  are shown in plot 14.1 and 14.2. This model capture really well the daily seasonality.

```{r LLT Seasonal Dummy}

mod2_ucm <- SSModel(train_ucm ~ SSMtrend(2, 
                  Q = list(matrix(NA), matrix(0))) +
                      SSMseasonal(24, 
                                  Q = matrix(NA),
                                  sea.type = 'dummy'),
                    H = matrix(NA))

fit2_ucm<- fitSSM(mod2_ucm,
                  inits=log(c(v_train_ucm/10,
                  v_train_ucm/100,
                  v_train_ucm/2,
                  v_train_ucm/50)))

#fit2_ucm$optim.out$convergence #convergencce

#-- previsoin
pred2_ucm <- predict(fit2_ucm$model,
                     n.ahead = length(validation_ucm)) 
pred2_ucm <- ts(pred2_ucm, start = 585, frequency = 24)


plot(validation_ucm, main = "14. LLT + daily s. dummy previsions", type = "l")+lines(pred2_ucm, col = "red", )+theme_gray()

#-- focus prima ed ultima settimana
plot(validation_ucm[1:(24*7)], type = "l", ylim = c(-2, 1), main = "14.1 Prevision LLT+ s. dummy first week", ylab='validation', xlab='time')+lines(pred2_ucm[1:(24*7),], col = "red")

plot(validation_ucm[(3504-24*7):3504], type = "l", ylim = c(-2, 1),main= "14.2 Prevision LLT+ s. dummy last week", ylab='validation', xlab='time')+lines(pred2_ucm[(3504-24*7):3504], col = "red")


mae2_ucm <- mean(abs(validation_ucm - pred2_ucm))
#mae2_ucm


```

To be sure to take in account all aspect of forecasting, a cyclic component is added to the prevoius model.
Considering that a cyclic pattern exists when data exhibit rises and falls that are not of fixed period and that the duration of these fluctuations is usually of at least 2 years. 
MAE of this model (0.5731635) increase compared to LLT+seasonal dummy model.

```{r LLT + season dummy + cycle (year)}
mod3_ucm <- SSModel(train_ucm ~ SSMtrend(2, Q = list(matrix(NA), matrix(0))) +
                      SSMseasonal(24, #-- period
                                  Q = matrix(NA),
                                  sea.type = 'dummy') + 
                      SSMcycle(24*365),
                    H = matrix(NA))

fit3_ucm<- fitSSM(mod3_ucm,
                  inits = log(c(v_train_ucm/10,
                                v_train_ucm/100,
                                v_train_ucm/2,
                                v_train_ucm/50)))
#fit3_ucm$optim.out$convergence #-- raggiunto num max di operazioni

#-- previsioni
pred3_ucm <- predict(fit3_ucm$model,
                     n.ahead = length(validation_ucm)) #-- Numerosità Validation (3504)
pred3_ucm <- ts(pred3_ucm, start = 585, frequency = 24)


plot(validation_ucm, main = "15. Prevision LLT+ s.dummy + cycle", type = "l", ylab='validation', xlab='time')
lines(pred3_ucm, col = "red")

#-- mae
mae3_ucm <- mean(abs(validation_ucm - pred3_ucm))
#mae3_ucm

```

Model with trigonometric seasonality is trained but it has no good result in term of MAE on validation set (MAE=0.6601218). 

```{r LLT,trigonometric season}
mod4_ucm <- SSModel(train_ucm ~ SSMtrend(2,
                                 Q = list(matrix(NA), matrix(0))) +
                      SSMseasonal(24, Q = matrix(NA),
                                  sea.type = "trigonometric"),
                    H = matrix(NA))

updt4 <- function(pars, model){
  model$Q[1, 1, 1] <- exp(pars[1])
  model$Q[2, 2, 1] <- exp(pars[2])
  model$Q[3, 3, 1] <- exp(pars[3])
  model$Q[4, 4, 1] <- exp(pars[4])
  diag(model$Q[5 : 25, 5 : 25, 1]) <- exp(pars[5])
  model$H[1, 1, 1] <- exp(pars[6])
  model
}

fit4_ucm <- fitSSM(mod4_ucm,
             log(c(v_train_ucm/10,
                   v_train_ucm/100,
                   v_train_ucm/75,
                   v_train_ucm/20,
                   v_train_ucm/50,
                   v_train_ucm/5)),
             updt4,
             control = list(maxit = 1000))

#fit4_ucm$optim.out$convergence 
#-- prevision
pred4_ucm <- predict(fit4_ucm$model,
                     n.ahead = length(validation_ucm)) 
pred3_ucm <- ts(pred4_ucm, start = 585, frequency = 24)

#-- mae 
mae4_ucm <- mean(abs(validation_ucm - pred4_ucm))
#mae4_ucm

```

Random Walk is than considered.

As previously seen this TS has a decreasing trend component due (probably) by covid effect on last month of training set (and also on validation).
Random walk is a substitute to this decreasing trend component. 
Focus plots for first week show a sovraestimation of values in first part of validation, but in the final part prevision are almost equal to true TS values.
MAE is 0.4814821.

```{r Random Walk + stag giornaliera + ciclo annuale}

mod5_ucm <- SSModel(train_ucm ~ SSMtrend(1, NA) +
                      SSMseasonal(24, NA, sea.type = "dummy") +
                    SSMcycle(24*365),
                    H = matrix(NA))
                
fit5_ucm<- fitSSM(mod5_ucm,
                  inits = log(c(v_train_ucm/10,
                                v_train_ucm/100,
                                v_train_ucm/2,
                                v_train_ucm/50)))

#fit5_ucm$optim.out$convergence

#-- previsions
pred5_ucm <- predict(fit5_ucm$model,
                     n.ahead = length(validation_ucm)) 
pred5_ucm <- ts(pred5_ucm, start = 585, frequency = 24)

plot(validation_ucm, main = "16. Previsions RW + s. dummy + cycle", type = "l")
lines(pred5_ucm, col = "red")

#-- focus prima ed ultima settimana
plot(validation_ucm[1:(24*7)], type = "l", ylim = c(-2, 1),main= "16.1 first week Previsions RW", ylab='validation', xlab='time')
lines(pred5_ucm[1:(24*7),], col = "red")

plot(validation_ucm[(3504-24*7):3504], type = "l", ylim = c(-2, 1),main= "16.2 last week Previsions RW", ylab='validation', xlab='time')
lines(pred5_ucm[(3504-24*7):3504], col = "red")

#-- mae
mae5_ucm <- mean(abs(validation_ucm - pred5_ucm))
#mae5_ucm



```

RW model is chosen as best model since its MAE on validation is the lowest of all UCM models trained.

```{r test prevision  with best UCM }
pred_best_ucm <- SSModel(ts_stand ~ SSMtrend(1, NA) +
                           SSMseasonal(24, NA, sea.type = "dummy") +
                           SSMcycle(24*365),
                         H = matrix(NA))

fit_best_ucm <- fitSSM(pred_best_ucm,
                  inits = log(c(v_train_ucm/10,
                                v_train_ucm/100,
                                v_train_ucm/2,
                                v_train_ucm/50)))

#fit_best_ucm$optim.out$convergence 
#-- prevision
pred_best_ucm <- predict(fit_best_ucm$model,
                     n.ahead = 1464) 
pred_best_ucm <- ts(pred_best_ucm, start = 731, frequency = 24)

plot(ts_stand, type = "l", xlim = c(0, 800), main='17. best UCM test prevision ', ylab='scaled TS')
lines(pred_best_ucm, col = "red")


#pred_test_ucm <- unscale(as.numeric(pred_best_ucm), ts_stand)
#pred_test_ucm <- ts(pred_test_ucm, start = 731, frequency = #24)

```

## ML models

### KNN

Training and validation set for KNN models are created respectively using  80% - 20% of data. 
Models are trined and evaluated on scaled data. 


2 type of KNN model are trained using first a week as
lenght of TS of which we need to find k  Neighbour, and in the second model using a day.

K parameter for first model (number of closest neighbour) is select in range 1-15, for second model the range is larger 1-20 (since lenght of p is smaller). Parameter tuning for k is made trought a cyclic function.
Horizon parameter is fixed to validation lenght.
Multiple-Step Ahead Strategy used is  recursive for both model and the combination function used to aggregate the targets associated with the nearest neighbors is the mean.

```{r training-validation an parameter KNN, echo=TRUE}
train_knn <- ts_stand[1:as.integer(nrow(ts)*0.8)]
validation_knn <-ts_stand[(1+as.integer(nrow(ts)*0.8)) : nrow(ts)]

#-- Parameter
#-- p is the lenght of TS of which we need to find k  Neighbour 
p <- 1:(24*7)
p1 <- 1:24

#-- k (Iperparameter k-nn)
k <- seq(1, 15, 2) 
k2 <-seq(1, 20, 2) 

#-- horzion of forecasat 
h <- 3504 

```

For model 1 with P=1 week, optimal number of NN is 9 with a corresponding MAE of 0.4345685.

For model 2 with P=1 day, optimal number of NN is 15 with a corresponding MAE of 0.4223891. 

```{r KNN algorithm}
for (el in k){
  mod_knn <- knn_forecasting(timeS = train_knn,
                       h = h, 
                       lags = p,
                       k = el,  
                       msas = "recursive", 
                       cf = "mean")

  print(el)
  pred_knn <- ts(mod_knn$prediction, start = 585, frequency = 24)
  mae_knn <- mean(abs(validation_knn - pred_knn))
  print(mae_knn)

}


for (el in k2){
  mod_knn <- knn_forecasting(timeS = train_knn,
                             h = h, 
                             lags = p1, 
                             k = el, 
                             msas = "recursive",
                             cf = "mean")           
  
  print(el)
  pred_knn <- ts(mod_knn$prediction, start = 585, frequency = 24)
  mae_knn <- mean(abs(validation_knn - pred_knn))
  print(mae_knn)
  
} #-- migliore: 15 knn con 0.4345685


mod_knn_best <- knn_forecasting(timeS = train_knn,
                             h = h, 
                             lags = p1, 
                             k = 15, 
                             msas = "recursive",
                             cf = "mean")           
  
pred_knn_best <- ts(mod_knn_best$prediction, start = 585, frequency = 24)
mae_knn_best <- mean(abs(validation_knn - pred_knn_best))
#mae_knn_best

prev_knn_best<-ts(pred_knn_best, start=14017)
autoplot(ts(ts_stand), main='18. KNN prevision on validation', ylab='ts_stand', xlab='time')+
autolayer(prev_knn_best)



```

### LSTM


Tensor Format:

Predictors (X) must be a 3D Array with dimensions: [samples, timesteps, features]: The first dimension is the length of values, the second is the number of time steps (lags), and the third is the number of predictors (1 if univariate or n if multivariate)
Outcomes/Targets (y) must be a 2D Array with dimensions: [samples, timesteps]: The first dimension is the length of values and the second is the number of time steps (lags)

Training/Testing dimension: 13652 1168

The training and testing length must be evenly divisible (e.g. training length / testing length must be a whole number)

Batch Size: 73 

The batch size is the number of training examples in one forward/backward pass of a RNN before a weight update
The batch size must be evenly divisible into both the training an testing lengths (e.g. training length / batch size and testing length / batch size must both be whole numbers)

Time Steps: 24*7

A time step is the number of lags included in the training/testing set

Epochs: 10

The epochs are the total number of forward/backward pass iterations
Typically more improves model performance unless overfitting occurs at which time the validation accuracy/loss will not improve.

```{r training-validation RNN, echo=TRUE}

datalags   <- 24*7 #-- Orizzonte di ricerca
batch.size <- 73 #-- prima era 146
epochs     <- 10


train_lstm <- as.data.frame(ts_stand[seq(16352 + datalags)])
validation_lstm <- as.data.frame(ts_stand[16352 + datalags + seq(1168 + datalags)])

df_val <- ts_stand[(17520-1167) : 17520,] #-- estrarre per mae

colnames(train_lstm) <- c("Valore")
colnames(validation_lstm) <- c("Valore")


x.train <- array(data = lag(train_lstm$Valore, datalags)[ -(1 : datalags)],
                 dim = c(nrow(train_lstm) - datalags, datalags, 1))

y.train <- array(data = train_lstm$Valore[-(1:datalags)],
                 dim = c(nrow(train_lstm)-datalags, 1))

x.test <- array(data = lag(validation_lstm$Valore,
                           datalags)[ -(1 : datalags)],
                dim = c(nrow(validation_lstm) - datalags, datalags, 1))

y.test <- array(data = validation_lstm$Valore[ -(1 : datalags)],
                dim = c(nrow(validation_lstm) - datalags, 1))

```

Long Short-Term Memory
A simple LSTM RNN is trained.
The model is composed of 2 layer: 

 - lstm layer with 16 neuron units
 - final dense layer 

This model obtain a MAE of  0.4086391
Plot of loss show that 10 epochs are sufficient to joint a minimum of 0,1 in loss function. 

```{r LSTM}

model_lstm1 <- keras_model_sequential() #-- Costruzione Strato per Strato

#--- 1. primo modello
model_lstm1 %>%
  layer_lstm(units = 16,                  #-- Numero Neuroni 
             input_shape = c(datalags, 1), #-- Dimensione Input
             batch_size = batch.size,
             return_sequences = TRUE,
             stateful = TRUE) %>%
  layer_dense(units = 1)                   #-- Layer Denso (Previsione)

model_lstm1 %>%
  compile(loss = 'mae',       #-- Funzione di Perdita
          optimizer = 'adam') #-- Ottimizzatore

print(model_lstm1)

#-- Addestramento
history_lstm1 <- model_lstm1 %>% fit(
  x = x.train, #-- Train
  y = y.train, #-- Test
  epochs = epochs, #-- Epoche di Addestramento
  batch_size = batch.size,
  verbose = 1,
  shuffle = FALSE
)
plot(history_lstm1, main='Loss of LSTM1')

pred_out_lstm1 <- model_lstm1 %>%
  predict(x.test,
          batch_size = batch.size) %>%
  .[, 168,1]

mae_lstm1 <- colMeans(drop_na(as.data.frame(abs(df_val - pred_out_lstm1))))
#mae_lstm1

plot(ts(ts_stand), type= "l")
lines(c(rep(NA, (16352 + datalags)), pred_out_lstm1),
      type = "l", lwd = 2, col = "red")

```

Another LSTM model is trained.

This model is composed of two LSTM layer, 2 layer of dropout that randomly sets input units to 0 with a frequency of rate 0.5  at each step during training time, which helps prevent overfitting, and a final dense layer.
To obtained good results 10 epochs are not sufficient, in fact with 10 epochs loss function is minimized at 0.2.
MAE for this model is  0.3982217.

```{r LSTM2}
#-- 2.. modello piï¿½ complesso
model_lstm2 <- keras_model_sequential() #-- Costruzione Strato per Strato

model_lstm2 %>%
  layer_lstm(units = 16,                   #-- Numero Neuroni 
             input_shape = c(datalags, 1), #-- Dimensione Input
             batch_size = batch.size,
             return_sequences = TRUE,
             stateful = TRUE) %>%          #-- Layer di tipo LSTM
  layer_dropout(rate = 0.5) %>%            #-- Layer di Dropout
  layer_lstm(units = 8,
             return_sequences = FALSE,
             stateful = TRUE) %>%          #-- Layer di tipo LSTM
  layer_dropout(rate = 0.5) %>%            #-- Layer di Dropout
  layer_dense(units = 1)                   #-- Layer Denso (Previsione)

model_lstm2 %>%
  compile(loss = 'mae',         #-- Funzione di Perdita
          optimizer = 'adam') #-- Ottimizzatore

model_lstm2

history_lstm2 <- model_lstm2 %>% fit(
  x = x.train, #-- Train
  y = y.train, #-- Test
  epochs = epochs, #-- Epoche di Addestramento
  batch_size = batch.size,
  verbose = 1,
  shuffle = FALSE
)
plot(history_lstm2, main='Loss of LSTM2')

pred_out_lstm2 <- model_lstm2 %>%
  predict(x.test,
          batch_size = batch.size)



mae_lstm2 <- colMeans(drop_na(as.data.frame(abs(df_val - pred_out_lstm2))))
#mae_lstm2


plot(ts(ts_stand), type= "l")
lines(c(rep(NA, (16352 + datalags)), pred_out_lstm2),
      type = "l", lwd = 2, col = "red")

```

 Gated Recurrent Units (GRU) model is trained.
 MAE for this model on validation set is 0.3938358.
 More epochs are needed to get better results, since the loss graph show a decreasing trend (and not a stationary situation).
```{r GRU}
mod_gru1 <- keras_model_sequential()

mod_gru1 %>%
  layer_gru(units = 20,
            input_shape = c(datalags, 1),
            batch_size  = batch.size,
            dropout = 0.3,
            recurrent_dropout = 0.5) %>%
  layer_dense(units = 1,
              activation = "linear")

mod_gru1 %>% 
  compile(loss = 'mae',
          optimizer = 'adam') #adam

mod_gru1

history_gru1 <- mod_gru1 %>% fit(
  x = x.train, #-- Train
  y = y.train, #-- Test
  epochs = epochs, #-- Epoche di Addestramento
  batch_size = batch.size,
  verbose = 1,
  shuffle = FALSE
)
plot(history_gru1, main='Loss of GRU')

pred_out_gru1 <- mod_gru1 %>%
  predict(x.test,
          batch_size = batch.size)

mae_gru1 <- colMeans(drop_na(as.data.frame(abs(df_val - pred_out_gru1))))
mae_gru1

plot(ts(ts_stand), type= "l")
lines(c(rep(NA, (16352 + datalags)), pred_out_gru1),
      type = "l", lwd = 2, col = "red")

```

LSTM1 model is chosen to perform prevision on test set. 
This is consider to be the best model because:

 - it is a simple model which obtain good result in terms of mae on valiadtion (0.40) compared to (0.39) of complex model.
 
 - in term of computational time 10 epochs are sufficient to reach a minimum in loss function, for the other 2 model more epochs are needed.
 
```{r best ml prevision}
data_to_csv <- read_csv2("C:\\Users\\davol\\Documents\\Magistrale\\time series\\PROJECT\\Previsioni_modelli.csv")
ts_ml<-ts(data_to_csv$ML, frequency=24, start=730)

autoplot(ts(ts$Valore, frequency = 24), main='Best ML model test prevision')+autolayer(ts_ml)
```

```{r test prevision with best ML model, eval=FALSE}
#-- modello migliore: lstm1

train_best_lstm <- as.data.frame(ts_stand[seq(16352 + datalags)])

x.train_best <- array(data = lag(train_lstm$Valore, datalags)
                      [ -(1 : datalags)],
                      dim = c(nrow(train_best_lstm)
                              - datalags, datalags, 1))

y.train_best <- array(data = train_lstm$Valore[-(1:datalags)],
                      dim = c(nrow(train_best_lstm)-datalags, 1))



#-- modello
model_best_ml <- keras_model_sequential() #-- Costruzione Strato per Strato

model_best_ml %>%
  layer_lstm(units = 16,                  #-- Numero Neuroni 
             input_shape = c(datalags, 1), #-- Dimensione Input
             batch_size = batch.size,
             return_sequences = TRUE,
             stateful = TRUE) %>%
  layer_dense(units = 1)                   #-- Layer Denso (Previsione)

model_best_ml %>%
  compile(loss = 'mae',       #-- Funzione di Perdita
          optimizer = 'adam') #-- Ottimizzatore

print(model_best_ml)

history_best <- model_best_ml %>% fit(
  x = x.train_best, #-- Train
  y = y.train_best, #-- Test
  epochs = 20, #-- Epoche di Addestramento
  batch_size = batch.size,
  verbose = 1,
  shuffle = FALSE
)
plot(history_best)


#-- function to predict TEST one step at the time
pred_list <- c(tail(ts_stand, n = datalags))

for (single_value in c(1: 1464)){
  single_value <- pred_list[(length(pred_list)-datalags):length(pred_list)]
  single_value <- array_reshape(single_value, c(1, 1, datalags))
  single_pred <- model_best_ml %>%
    predict(single_value)
  
  print(single_pred)
  pred_list <- append(single_pred)
}  


plot(ts(ts_stand), type= "l", xlim = c(0, 18984))
lines(c(rep(NA, 17520), pred_out_best_ml[,1,1][1:(61*24)]),
      type = "l", lwd = 2, col = "red")
```
 
## Conclusion

Best results in term of MAE on validation set are obtained by ML model and Arimax model.

In particular it has been shown that simple arima model and UCM model are affected by the (strong) decreasing trend. In this situation regressors are necessary to model yearly seasonality. 
UCM models are the worst of all trained model. Probably more complex model should have been trained. 
Moreover to solve Covid problem a level shift could be added in the models.




```{r scrittura csv, eval=FALSE}

df_date <- c()
df_hours <- c()
for (j in c(1:61)){
  for (i in c(1:24)){
    
    date <- as_date("2020-09-01") + j
    hours <- i
    
    df_date <- append(df_date, date)
    df_hours <- append(df_hours, hours)
  }
}

data <- cbind(as.Date(df_date),
              df_hours,
              unscale(pred_best_arima$mean, ts_stand),
              unscale(pred_best_ucm, ts_stand))
              #unscale(pred_out_best_ml[, 1, 1][1:(61*24)], ts_stand))
colnames(data) <- c("Data", "Ora", "ARIMA", "UCM")

write_csv2(as.data.frame(data), "calendar.csv")


data_to_csv <- read_csv2("Previsioni_modelli.csv")

#plot(ts$Valore, type= "l", xlim = c(0, 20000))
#lines(c(rep(NA, 17520),
#            unscale(pred_out_best_ml[, 1, 1][1:(61*24)], ts_stand)),
#      type = "l", lwd = 2, col = "red")    

```

